A crop modelling strategy to improve cacao quality and productivity

### 2021-12-16 paperModel/paperCacaoModel_round1.tex
correction round 1

Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly. The farmer practice of empirically harvesting from 5 to 6 months or 180 Days After Flowering  (DAF) date, produced a mix of quality of cacao beans that are then fermented. \textcolor{blue}{The problem is that  the seeds are often  unripe or  over ripe germinating inside the pod affecting the cocoa quality}. This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the final cocoa product \citep{Escobar2021}.  \textcolor{blue}{Therefore,  It is important to identify the best moment to harvest  considering physiological responses affected by  soil types , cultivar specifications and weather variables such as rain, solar radiation and wind.  Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits, hence it will improve the quality of seeds.}

@@

\subsection{Cocoa Crop Model Simulations}
Although cocoa is a relevant crop and there is an extensive agronomic literature, there is only one physiological crop model specific for cocoa so far. The (SUCROS-Cocoa) developed by \cite{zuidema2005}. However, the code was not easy available for adaptations. In contrast, the SIMPLE model has an open code in R, which we could adapt such a model would be very useful to compare yields and predict harvest date in different climates. As the harvest day was predicted from the FD (Fig. \ref{fig:dayH}, consequently, biomass production and fSolar presented a crop cycle shorter than 180 DAF (Fig. \ref{fig:m1}, a and b). These simulations are coincident with results presented by \cite{lopez2018}, where physiological maturity of coca pod varies from 140 to 162 DAF.  Our results showed how the harvest day can vary depending on the accumulate temperature during each specific crop cycle simulated (Fig. \ref{fig:dayH}). \textcolor{blue} {Harvesting cocoa fruits at the right moment the quality of the seed inside the pod can improve avoiding the germination before collect the pot from  cacao trees.}\textcolor{olive} {An research made by Gallego (2021), cacao clones CCN51 and ICS95 were handed-pollinated and pods were collected at 6, 7, and 8 months of ripeness after pollination. Cacao fruits were harvested from the municipality of Belalcazar, Caldas, Colombia to analyse metabolites related with cocoa quality were more expressed when fruits are collected at 6, 7, and 8 months after flowering.  Results showed that the metabolites that mostly explain the differences between the classes ( metabolites with the highest VIP) are highly overexpressed at month 6 and then decline or under-expressed at the next months.  No significant differences were found in their metabolic profile Between month 7 and 8.
Quality of cocoa is defined by  some metabolites that are associated with flavonoids or colouration. When the expression of flavonoids decrease , indicates  that  metabolites  associated with flavonoids  are being oxidized and metabolites are converted by cells into other less oxidizing metabolites. The colouration does not show a greater difference after month 6. If there is no good  accumulation of flavonoids which are also related with metabolites associated with colouration, the quality of cocoa seed decrease. Ripening is associated with changes in colouration and these in turn with flavonoids so month 6 was ideal for harvesting assuring the high quality of cocoa. } \textcolor{blue} { Therefore, the cacao model can predict with more detail the optimum day to harvest the pod ensuring the quality of the cacao bean  destined for Fine-cocoa products. }

>>>>

Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly. The farmer practice of empirically harvesting from 5 to 6 months or 180 Days After Flowering  (DAF) date, produced a mix of quality of cacao beans that are then fermented. \textcolor{blue}{The problem is that  the seeds are often  unripe or  over ripe germinating inside the pod affecting the cocoa quality}. This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the final cocoa product \citep{Escobar2021}.  \textcolor{blue}{Therefore,  It is important to identify the best moment to harvest  considering physiological responses affected by  soil types , cultivar specifications and weather variables such as rain, solar radiation and wind.  Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits, hence it will improve the quality of seeds.  }

\textcolor{blue}{Climate change could severely impact cacao production as cacao crop is often described as being sensitive to climate change  producing negatively impact its growth, productivity and yield quality  \citep{zuidema2005, Ceccarelli2021}. However, there are some   	 adaptive responses to mitigate negative climate  effects, for example elevated CO$_{2}$ and tolerant varieties \citep{lahive2019, Ceccarelli2021}.  The role of genetic diversity in producing more resilient cacao  to drought and temperature tolerance, or increased CO2 response,, are presented in \citep{Medina2017} .Thus crop models that include weather data can provide a tool to study the impact of climate change over the quality of cacao. }

@@

\subsection{Cocoa Crop Model Simulations}
Although cocoa is a relevant crop and there is an extensive agronomic literature, there is only one physiological crop model specific for cocoa so far is The (SUCROS-Cocoa) developed by \cite{zuidema2005}. However, the code was not easy available for adaptations. In contrast, the SIMPLE model has an open code in R, which we could adapt such a model would be very useful to compare yields and predict harvest date in different climates. As the harvest day was predicted from the FD (Fig. \ref{fig:dayH}, consequently, biomass production and fSolar presented a crop cycle shorter than 180 DAF (Fig. \ref{fig:m1}, a and b). These simulations are coincident with results presented by \cite{lopez2018}, where physiological maturity of coca pod varies from 140 to 162 DAF.  Our results showed how the harvest day can vary depending on the accumulate temperature during each specific crop cycle simulated (Fig. \ref{fig:dayH}). \textcolor{blue} {Harvesting cocoa fruits at the right moment the quality of the seed inside the pod can improve avoiding the germination before collect the pod from  cacao trees.  Therefore, the cacao model can predict with more detail the optimum day to harvest the pod ensuring the quality of the cacao bean  destined for Fine-cocoa products.}
\textcolor{olive} {Fruit ripening is a highly coordinated developmental process that coincides with seed maturation \citep{Osorio2013} . As cacao is a non climacteric fruit, the pod needs be connected to the mother plant until maturation. Therefore, an indirect way to infer the seed quality is checking the pod quality. Pod quality can be determined as the physico-chemical changes occurring at the peel level, (including colouration changes, and chemical dynamical changes in metabolites production such as sugars, phenolics, fatty acids) which give information about the seed status. For instance, clear evidence of above occurs when cacao ripe fruits stay for a longer time on the tree without be harvested on the right moment, causing that seeds germinate  inside the pod and damaging the cacao production for high-quality taste and affecting the final yield. 
An ongoing project conducted by BIOS is showing us the expression patterns for metabolites associated with cacao ripening. In this case, we found sugars, phenolic, fatty acids and flavonoids which according to the ripening stage change their expression.  Furthermore, we are identifying biochemical markers of fruit ripening. We are also connecting these chemical changes with seed quality. It let us understand how the ripen occur in cacao and suggest the best moment to harvest.}
